---
title: "privateEC Quick Start"
author: "Bill C. White"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Privacy preserving Evaporative Cooling feature selection and classification with Relief-F and Random Forests

Methods are described in the following publication.

Trang Le, W. K. Simmons, M. Misaki, B.C. White, J. Savitz, J. Bodurka, and B. A. McKinney. “Privacy preserving evaporative cooling feature selection and classification with Relief-F and Random Forests,” Bioinformatics. Accepted. [Bioinformatics Abstract](https://doi.org/10.1093/bioinformatics/btx298). 2017.

## Example Run and Results

```{r}
library(privateEC)
num.samples <- 100
num.variables <- 100
pb <- 0.1
nbias <- pb * num.variables
signals <- sprintf("gene%04d", 1:nbias)
sim.data <- createSimulation(d=num.variables, n=num.samples,
                             type="sva", verbose=FALSE)
pec.results <- privateEC(data.sets=sim.data, is.simulated=TRUE, n=num.samples,
                         signal.names=signals, verbose=FALSE, update.freq=5)
```

```{r, echo=FALSE}
knitr::kable(pec.results$plots.data, caption="Algorithm Iterations",
             row.names=FALSE)
```

```{r, echo=FALSE, fig.width=6, fig.align='center'}
library(ggplot2)
ggplot(pec.results$melted.data, aes(x=num.atts, y=value, colour=variable)) +
  geom_point(size=1) + geom_line()
```
