---
title: "Real Data Example"
author: "Bill C. White"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Real Data Example}
  %\VignetteEncoding{UTF-8}
---

# Privacy preserving Evaporative Cooling feature selection and classification with Relief-F and Random Forests

Methods are described in the following publication.

Trang Le, W. K. Simmons, M. Misaki, B.C. White, J. Savitz, J. Bodurka, and B. A. McKinney. “Differential privacy-based Evaporative Cooling feature selection and classification with Relief-F and Random Forests,” Bioinformatics. Accepted. [Bioinformatics Abstract](https://doi.org/10.1093/bioinformatics/btx298). 2017.

## Load a subset of the resting state fMRI region correlation

```{r}
library(privateEC)
data(rsfMRIcorrMDD)
# ~100 variables for a test
data.width <- ncol(rsfMRIcorrMDD)
real.data.sets <- splitDataset(all.data = rsfMRIcorrMDD[, (data.width - 101):data.width],
                               pct.train = 0.5,
                               pct.holdout = 0.5,
                               pct.validation = 0,
                               label = "class")
real.result <- privateEC(train.ds = real.data.sets$train,
                         holdout.ds = real.data.sets$holdout,
                         validation.ds = NULL,
                         label = "class",
                         is.simulated = FALSE,
                         update.freq = 5,
                         verbose = FALSE)
```

```{r, echo=FALSE}
knitr::kable(real.result$algo.acc, 
             caption = "Algorithm Iterations",
             row.names = FALSE, 
             digits = 3)
```

```{r, echo=FALSE, fig.width=7, fig.width=7, fig.align='center'}
plotRunResults(real.result)
```
