---
title: "Example1"
author: "Bill C. White"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
  
vignette: >
  %\VignetteIndexEntry{Example1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Privacy preserving Evaporative Cooling feature selection and classification with Relief-F and Random Forests

Methods are described in the following publication.

Trang Le, W. K. Simmons, M. Misaki, B.C. White, J. Savitz, J. Bodurka, and B. A. McKinney. “Privacy preserving evaporative cooling feature selection and classification with Relief-F and Random Forests,” Bioinformatics. Accepted. [Bioinformatics Abstract](https://doi.org/10.1093/bioinformatics/btx298). 2017.

## Example Run and Results

```{r}
library(privateEC)
myrun <- "001"
n <- 100
num.vars <- 100
bias <- 0.4
type <- "mainEffect"
pct.signals <- 0.1
update.freq <- 5
verbose <- FALSE
shortname <- "Example1"

data.sets <- createSimulation(n=n,
                              num.vars=num.vars,
                              pct.signals=pct.signals,
                              bias=bias,
                              shortname=shortname,
                              sim.type=type,
                              myrun=myrun,
                              verbose=verbose,
                              save.file=FALSE)

pec.result <- privateEC(train.ds=data.sets$train,
                        holdout.ds=data.sets$holdout,
                        validation.ds=data.sets$validation,
                        label=data.sets$class.label,
                        is.simulated=TRUE,
                        shortname=shortname,
                        bias=bias,
                        myrun=myrun,
                        update.freq=update.freq,
                        save.file=NULL,
                        verbose=verbose,
                        signal.names=data.sets$signal.names)
```

```{r, echo=FALSE}
knitr::kable(pec.result$algo.acc, caption="Algorithm Iterations",
             row.names=FALSE, digits=3)
```

```{r, echo=FALSE, fig.width=14, fig.width=7, fig.align='center'}
# library(ggplot2)
# ggplot(pec.result$melted.data, aes(x=num.atts, y=value, colour=variable)) +
#   geom_point(size=1) + geom_line()
plot(pec.result$algo.acc$vars.remain, 
     pec.result$algo.acc$holdout.acc, 
     col="red", pch=16, type='b', cex=0.75, 
     main="One run of privateEC",
     xlab="Number of Attributes in Model",
     ylab="Accuracy")
points(pec.result$algo.acc$vars.remain, 
       pec.result$algo.acc$train.acc, 
       col="green", pch=1, type='b', cex=0.75)
points(pec.result$algo.acc$vars.remain, 
       pec.result$algo.acc$validation.acc, 
       col="blue", pch=4, type='b', cex=0.75)
legend("topright", c("Train", "Holdout", "Test"), 
       pch=c(16, 1, 4), col=c("red", "green", "blue"), cex=0.75)
```
