---
title: "pEC_nCV_Vignette"
author: Saeid Parvandeh
date: "`r Sys.Date()`"
output:
  github_document:
  keep_md: yes
---
## Privacy preserving Evaporative Cooling (PrivateEC), Consensus Nested Cross Validation (cnCV), and Regular Nested Cross Validation (rnCV) for feature selection and classification with Relief-F and Random Forests

PrivateEC methods are described in the following publication.

Trang T. Le, W. K. Simmons, M. Misaki, B.C. White, J. Savitz, J. Bodurka, and B. A. McKinney. "Differential privacy-based evaporative cooling feature selection and classification with Relief-F and Random Forests," <i>Bioinformatics</i>, Volume 33, Issue 18, 15 September 2017, Pages 2906–2913. [free](https://academic.oup.com/bioinformatics/article/33/18/2906/3796394/Differential-privacybased-evaporative-cooling?guestAccessKey=b91832c8-6034-46d5-95ce-382fc71c1df0)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(privateEC)
```

## Simulation example comparing private Evaporative Cooling (pEC) with standard nested cross-validation (nCV) and a new consensus-feature nested cross validation (cnCV)

### Simulate Balanced Case-Control Data with Numeric Predictors with Main Effects
```{r, message=FALSE}
library(privateEC)
n.samples <- 300     # 100 samples in train/holdout/test
n.variables <- 100   # 100 features
label <- "class"
type <- "mainEffect" # main effect simulatios
bias <- 0.6          # moderate effect size
pct.signals <- 0.1   # pct functional features
verbose <- FALSE

data.sets <- createSimulation(num.samples = n.samples,
                              num.variables = n.variables,
                              pct.signals = pct.signals,
                              label = label,
                              bias = bias,
                              pct.train = 1/3,
                              pct.holdout = 1/3,
                              pct.validation = 1/3,
                              sim.type = type,
                              save.file = NULL,
                              verbose = verbose)
```

### Run pEC on simulated data

```{r, message=FALSE}
update.freq <- 5   # attributes evaporated per iteration
num_tree <- 500
importance.algorithm <- "ReliefFequalK"

pec.result <- privateEC(train.ds = data.sets$train,                   # training split
                        holdout.ds = data.sets$holdout,               # holdout split
                        validation.ds = data.sets$validation,         # independent validation set
                        label = data.sets$label,                      # class label
                        importance.algorithm = importance.algorithm,  # feature selection method (ReliefF)
                        relief.k.method = "k_half_sigma",             # ReliefF nearest neighbor k estimate
                        rf.ntree = num_tree,                          # random forest classifier
                        is.simulated = TRUE,                          # indicates there will be a validation set
                        bias = bias,                                  # pEC privacy temperature noise
                        update.freq = update.freq,                    # number of attributes to evaporate per step
                        save.file = NULL,                             # save results
                        signal.names = data.sets$signal.names,        # functional variable names
                        use.nestedCV = F,
                        verbose = verbose)
```

### pEC results
#### Table of iterations
```{r, echo=FALSE}
knitr::kable(pec.result$algo.acc, caption = "Evaporation Iterations",
             row.names = FALSE, digits = 3)
```

#### Plot of results
```{r, echo=FALSE, fig.width=7, fig.width=7, fig.align='center'}
plotRunResults(pec.result)
```

#### pEC selected features
```{r, echo=T}
cat("\n Max Holdout Accuracy Step [",which.max(pec.result$algo.acc$holdout.acc),"]\n")
cat("\n Accuracies: ")
print(pec.result$algo.acc[which.max(pec.result$algo.acc$holdout.acc),])
cat("\n Selected Features \n [",pec.result$atts.remain[[which.max(pec.result$algo.acc$holdout.acc)]],"]\n")
```

### Run regular nested CV

```{r, message=FALSE}
rncv_result <- regular_nestedCV(train.ds = data.sets$train, 
                                validation.ds =  data.sets$holdout, 
                                label = data.sets$label,
                                is.simulated = TRUE,
                                ncv_folds = c(10, 10),
                                param.tune = FALSE,
                                learning_method = "rf", 
                                importance.algorithm = importance.algorithm,
                                num_tree = num_tree,
                                verbose = verbose)
```

### rnCV results
```{r, message=FALSE, echo=T}
cat("\n Train Accuracy [",rncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",rncv_result$Validation,"]\n")
cat("\n Selected Features \n [",rncv_result$Features,"]\n")
cat("\n Elapsed Time [",rncv_result$Elapsed,"]\n")
```

### Run consensus nested CV
```{r, message=FALSE}
cncv_result <- consensus_nestedCV(train.ds = data.sets$train, 
                                  validation.ds =  data.sets$holdout, 
                                  label = data.sets$label,
                                  is.simulated = TRUE,
                                  ncv_folds = c(10, 10),
                                  param.tune = FALSE,
                                  learning_method = "rf", 
                                  importance.algorithm = importance.algorithm,
                                  num_tree = num_tree,
                                  verbose = verbose)
```

### cnCV results

```{r, echo=T}
cat("\n Train Accuracy [",cncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",cncv_result$Validation,"]\n")
cat("\n Selected Features \n [",cncv_result$Features,"]\n")
cat("\n Elapsed Time [",cncv_result$Elapsed,"]\n")
```

### Simulate Imbalanced Data (parameter in effect: pct.imbalance, where 0/1 means all ctrls/cases)
```{r, message=FALSE}
n.samples <- 100
n.variables <- 100
label <- "class"
pct.imbalance = 0.56
type <- "mainEffect"
importance.algorithm <- "ReliefFequalK"
bias <- 0.4
pct.signals <- 0.1
update.freq <- 5
num_tree <- 500
verbose <- FALSE

data.sets <- createSimulation(num.samples = n.samples,
                              num.variables = n.variables,
                              pct.signals = pct.signals,
                              label = label,
                              pct.imbalance = pct.imbalance,
                              bias = bias,
                              pct.train = 1/3,
                              pct.holdout = 1/3,
                              pct.validation = 1/3,
                              sim.type = type,
                              save.file = NULL,
                              verbose = verbose)
```

### Run pEC on imbalanced simulated data

```{r, message=FALSE}
pec.result <- privateEC(train.ds = data.sets$train,
                        holdout.ds = data.sets$holdout,
                        validation.ds = data.sets$validation,
                        label = data.sets$label,
                        importance.algorithm = importance.algorithm,
                        rf.ntree = num_tree,
                        is.simulated = TRUE,
                        bias = bias,
                        update.freq = update.freq,
                        save.file = NULL,
                        signal.names = data.sets$signal.names,
                        verbose = verbose)
```

### pEC results
#### Table of iterations
```{r, echo=FALSE}
knitr::kable(pec.result$algo.acc, caption = "Algorithm Iterations",
             row.names = FALSE, digits = 3)
```

#### Plot of results
```{r, echo=FALSE, fig.width=7, fig.width=7, fig.align='center'}
plotRunResults(pec.result)
```

#### pEC selected features
```{r, echo=FALSE}
cat("\n Selected Features \n [",pec.result$atts.remain[[which.max(pec.result$algo.acc$holdout.acc)]],"]\n")
```

### Run regular nested CV

```{r, message=FALSE}
rncv_result <- regular_nestedCV(train.ds = data.sets$train, 
                                validation.ds =  data.sets$holdout, 
                                label = data.sets$label,
                                is.simulated = TRUE,
                                ncv_folds = c(10, 10),
                                param.tune = FALSE,
                                learning_method = "rf", 
                                importance.algorithm = importance.algorithm,
                                num_tree = num_tree,
                                verbose = verbose)
```

### rnCV results
```{r, message=FALSE, echo=FALSE}
cat("\n Train Accuracy [",rncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",rncv_result$Validation,"]\n")
cat("\n Selected Features \n [",rncv_result$Features,"]\n")
cat("\n Elapsed Time [",rncv_result$Elapsed,"]\n")
```

### Run consensus nested CV
```{r, message=FALSE}
cncv_result <- consensus_nestedCV(train.ds = data.sets$train, 
                                  validation.ds =  data.sets$holdout, 
                                  label = data.sets$label,
                                  is.simulated = TRUE,
                                  ncv_folds = c(10, 10),
                                  param.tune = FALSE,
                                  learning_method = "rf", 
                                  importance.algorithm = importance.algorithm,
                                  num_tree = num_tree,
                                  verbose = verbose)
```

### cnCV results

```{r, echo=FALSE}
cat("\n Train Accuracy [",cncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",cncv_result$Validation,"]\n")
cat("\n Selected Features \n [",cncv_result$Features,"]\n")
cat("\n Elapsed Time [",cncv_result$Elapsed,"]\n")
```

### Simulate Mix Data (parameter in effect: pct.imbalance, where 0/1 means all ctrls/cases, pct.mixed)
```{r, message=FALSE}
n.samples <- 100
n.variables <- 100
label <- "class"
pct.imbalance = 0.5
pct.mixed = 0.5
mixed.type <- c("mainEffect", "interactionScalefree")
importance.algorithm <- "ReliefFequalK"
bias <- 0.4
pct.signals <- 0.1
update.freq <- 5
num_tree <- 500
verbose <- FALSE

data.sets <- createMixedSimulation(num.samples = n.samples,
                                   num.variables = n.variables,
                                   pct.signals = pct.signals,
                                   label = label,
                                   pct.imbalance = pct.imbalance,
                                   pct.mixed = pct.mixed,
                                   mixed.type = mixed.type,
                                   bias = bias,
                                   pct.train = 1/3,
                                   pct.holdout = 1/3,
                                   pct.validation = 1/3,
                                   save.file = NULL,
                                   verbose = verbose)
```

### Run pEC on mixed simulated data

```{r, message=FALSE}
pec.result <- privateEC(train.ds = data.sets$train,
                        holdout.ds = data.sets$holdout,
                        validation.ds = data.sets$validation,
                        label = data.sets$label,
                        importance.algorithm = importance.algorithm,
                        rf.ntree = num_tree,
                        is.simulated = TRUE,
                        bias = bias,
                        update.freq = update.freq,
                        save.file = NULL,
                        signal.names = data.sets$signal.names,
                        verbose = verbose)
```

### pEC results
#### Table of iterations
```{r, echo=FALSE}
knitr::kable(pec.result$algo.acc, caption = "Algorithm Iterations",
             row.names = FALSE, digits = 3)
```

#### Plot of results
```{r, echo=FALSE, fig.width=7, fig.width=7, fig.align='center'}
plotRunResults(pec.result)
```

#### pEC selected features
```{r, echo=FALSE}
cat("\n Selected Features \n [",pec.result$atts.remain[[which.max(pec.result$algo.acc$holdout.acc)]],"]\n")
```

### Run regular nested CV

```{r, message=FALSE}
rncv_result <- regular_nestedCV(train.ds = data.sets$train, 
                                validation.ds =  data.sets$holdout, 
                                label = data.sets$label,
                                is.simulated = TRUE,
                                ncv_folds = c(10, 10),
                                param.tune = FALSE,
                                learning_method = "rf", 
                                importance.algorithm = importance.algorithm,
                                num_tree = num_tree,
                                verbose = verbose)
```

### rnCV results
```{r, message=FALSE, echo=FALSE}
cat("\n Train Accuracy [",rncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",rncv_result$Validation,"]\n")
cat("\n Selected Features \n [",rncv_result$Features,"]\n")
cat("\n Elapsed Time [",rncv_result$Elapsed,"]\n")
```

### Run consensus nested CV
```{r, message=FALSE}
cncv_result <- consensus_nestedCV(train.ds = data.sets$train, 
                                  validation.ds =  data.sets$holdout, 
                                  label = data.sets$label,
                                  is.simulated = TRUE,
                                  ncv_folds = c(10, 10),
                                  param.tune = FALSE,
                                  learning_method = "rf", 
                                  importance.algorithm = importance.algorithm,
                                  num_tree = num_tree,
                                  verbose = verbose)
```

### cnCV results

```{r, echo=FALSE}
cat("\n Train Accuracy [",cncv_result$cv.acc,"]\n")
cat("\n Validation Accuracy [",cncv_result$Validation,"]\n")
cat("\n Selected Features \n [",cncv_result$Features,"]\n")
cat("\n Elapsed Time [",cncv_result$Elapsed,"]\n")
```

### Simulate Quantitative Outcome Data (parameter in effect: label = "qtrait", type = "mainEffect", importance.algorithm = "RReliefFequalK")
```{r, message=FALSE}
n.samples <- 100
n.variables <- 100
label <- "qtrait"
type <- "mainEffect"
importance.algorithm <- "RReliefFequalK"
bias <- 0.4
pct.signals <- 0.1
update.freq <- 5
num_tree <- 500
verbose <- FALSE

data.sets <- createSimulation(num.samples = n.samples,
                              num.variables = n.variables,
                              pct.signals = pct.signals,
                              label = label,
                              bias = bias,
                              pct.train = 1/3,
                              pct.holdout = 1/3,
                              pct.validation = 1/3,
                              sim.type = type,
                              save.file = NULL,
                              verbose = verbose)
```

### Run pEC on quantitative simulated data

```{r, message=FALSE}
pec.result <- privateEC(train.ds = data.sets$train,
                        holdout.ds = data.sets$holdout,
                        validation.ds = data.sets$validation,
                        label = data.sets$label,
                        importance.algorithm = importance.algorithm,
                        rf.ntree = num_tree,
                        is.simulated = TRUE,
                        bias = bias,
                        update.freq = update.freq,
                        save.file = NULL,
                        signal.names = data.sets$signal.names,
                        verbose = verbose)
```

### pEC results
#### Table of iterations
```{r, echo=FALSE}
knitr::kable(pec.result$algo.acc, caption = "Algorithm Iterations",
             row.names = FALSE, digits = 3)
```

#### Plot of results
```{r, echo=FALSE, fig.width=7, fig.width=7, fig.align='center'}
plotRunResults(pec.result)
```

#### pEC selected features
```{r, echo=FALSE}
cat("\n Selected Features \n [",pec.result$atts.remain[[which.max(pec.result$algo.acc$holdout.acc)]],"]\n")
```

### Run regular nested CV

```{r, message=FALSE}
rncv_result <- regular_nestedCV(train.ds = data.sets$train, 
                                validation.ds =  data.sets$holdout, 
                                label = data.sets$label,
                                is.simulated = TRUE,
                                ncv_folds = c(10, 10),
                                param.tune = FALSE,
                                learning_method = "rf", 
                                importance.algorithm = importance.algorithm,
                                num_tree = num_tree,
                                verbose = verbose)
```

### rnCV results
```{r, echo=FALSE}
cat("\n Train R-squared [",rncv_result$cv.acc,"]\n")
cat("\n Validation R-squared [",rncv_result$Validation,"]\n")
cat("\n Selected Features \n [",rncv_result$Features,"]\n")
cat("\n Elapsed Time [",rncv_result$Elapsed,"]\n")
```

### Run consensus nested CV
```{r, message=FALSE}
cncv_result <- consensus_nestedCV(train.ds =  data.sets$train, 
                                  validation.ds =  data.sets$holdou, 
                                  label = data.sets$label,
                                  is.simulated = TRUE,
                                  ncv_folds = c(10, 10),
                                  param.tune = FALSE,
                                  learning_method = "rf", 
                                  importance.algorithm = importance.algorithm,
                                  num_tree = num_tree,
                                  verbose = verbose)
```

### cnCV results

```{r, echo=FALSE}
cat("\n Train R-squared [",cncv_result$cv.acc,"]\n")
cat("\n Validation R-squared [",cncv_result$Validation,"]\n")
cat("\n Selected Features \n [",cncv_result$Features,"]\n")
cat("\n Elapsed Time [",cncv_result$Elapsed,"]\n")
```


